---
title: "Info 201 Project"
output: html_notebook
---

# Description

Influenza vaccine data source: <https://data.cdc.gov/Flu-Vaccinations/Influenza-Vaccination-Coverage-for-All-Ages-6-Mont/vh55-3he6> covid vaccine data source: <https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh> Question: Does the % of influenza vaccinations pre covid (2019-2020) correlate to the % of Covid-19 Dose 1 Vaccinations (as of 2023-05-10) *May add more time stamps* in US Counties?

Theory: We may be able to extrapolate based on our question whether or not people's willingness to take vaccines in general (the assumption is that influenza vaccines can represent "vaccines in general") impacts their willingness to take the Covid-19 vaccine (and if so, to what extent?).


# Load libraries and data, print samples

```{r}
library(tidyverse)
influenza_vaccine_data = read.csv("data/Influenza_data.csv")
print(sample_n(influenza_vaccine_data,10))
covid_vaccine_data = read.csv("data/Covid_data.csv") # Note: the covid data is huge relative to the influenza data, so it might take some time to load
print(sample_n(covid_vaccine_data,10))


```

# Data Cleaning
Filter out irrelevant geography types to get only the counties, clean up some other aspects. It is important to point out that this data is estimated based off of state data and population distribution, meaning the actual percentage of each county may not be accurate. Unfortunately, we do not have the same amount of precision as we do with the covid data as there were less resources dedicated to influenza. We are only focusing on 2019-2020 influenza data because we just need a snapshot of how willing people are to taking vaccines shortly before COVID-19.

```{r}
covid_vaccine_data_cleaned = covid_vaccine_data %>%
  select(FIPS, Administered_Dose1_Pop_Pct, Date, Census2019_18PlusPop) %>% # Remove unnecessary columns. We'll only look at dose1 in the two dose covid vaccine.
  #mutate(Recip_County = sub(" .*", "", Recip_County)) %>% # Remove the unnecessary "county" specification after every county name
  mutate(Date = as.Date(Date,format = "%m/%d/%Y")) %>% # Convert dates into a R friendly format
  filter(!is.na(as.numeric(FIPS))) %>% # Remove unique region-county identifiers
  rename(Covid_Percentage = Administered_Dose1_Pop_Pct) %>%
  mutate(FIPS = as.numeric(FIPS))

# Note: This dataset has column names which haves spaces in them. When refering to these columns, enclose the column name with backticks (`). FIPS = County identifier (Since many counties have duplicate names).
influenza_vaccine_data_cleaned = influenza_vaccine_data %>%
  filter(Vaccine == "Seasonal Influenza") %>% # Only the general influenza vaccine will be considered as the special variants make up a negligible proportion
  filter(`Geography.Type` == "Counties") %>% # Filter out non-county data (state/local areas)
  filter(`Dimension.Type` == "Age") %>% # Filter out vaccine demographic data (data only includes people >=18)
  filter(!is.na(as.numeric(FIPS))) %>% # Remove unique region-county identifiers
  filter(`Season.Survey.Year` == "2019-20") %>% # Only keep 2019-2020 data for simplicity (may add more dates later)
  rename(Influenza_Percentage = `Estimate....`) %>%
  mutate(Influenza_Percentage = as.numeric(Influenza_Percentage)) %>%# Convert the estimate % from character to numeric
  #rename(Date = "Season/Survey Year")
  select(FIPS, Influenza_Percentage) # Filter out irrelevant columns

print(influenza_vaccine_data_cleaned)
print(covid_vaccine_data_cleaned)

```

# Merging dataframes
```{r}
#merging
joinedData <- inner_join(influenza_vaccine_data_cleaned, covid_vaccine_data_cleaned, by = "FIPS")
#print(joinedData)

FirstMonth <- joinedData%>%
  select(Covid_Percentage, Date,FIPS, Census2019_18PlusPop )%>%
  filter(Date == "2021-01-17")%>%
  rename(fips = FIPS)

SixMonths <- joinedData %>%
  select(Covid_Percentage, Date,FIPS, Census2019_18PlusPop)%>%
  filter(Date ==  "2021-06-17")%>%
  rename(fips = FIPS)

OneYear <- joinedData %>%
  select(Covid_Percentage, Date,FIPS, Census2019_18PlusPop)%>%
  filter(Date == "2022-01-17")%>%
  rename(fips = FIPS)

influenza_vaccine_data_cleaned1 <- influenza_vaccine_data_cleaned%>%
  rename(fips = FIPS)

FirstMonthjoinedPlus <- inner_join(FirstMonth, influenza_vaccine_data_cleaned1, by = "fips" )
SixMonthsjoinedPlus <- inner_join(SixMonths, influenza_vaccine_data_cleaned1, by = "fips" )
oneYearjoinedPlus <- inner_join(OneYear, influenza_vaccine_data_cleaned1, by = "fips" )
print(FirstMonthjoinedPlus)
```
# Plotting
This data is severely skewed by outliers
```{r}
#FirstMonthJoined
#SixMonthsjoined
#OneYearJoined


FirstMonthjoinedPlus %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage)) +
        geom_point()+
        labs(title = "One months after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

 # cor(FirstMonthjoinedPlus$Influenza_Percentage, FirstMonthjoinedPlus$Covid_Percentage)


SixMonthsjoinedPlus %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage))+
        geom_point()+
        labs(title = "Six months after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

##  cor(SixMonthsCors$Influenza_Percentage, SixMonthsCors$Covid_Percentage)

oneYearjoinedPlus %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage)) +
        geom_point()+
        labs(title = "One Year after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

#OneYearCors<- oneYearjoinedPlus%>%
  #cor(Influenza_Percentage, Covid_Percentage)
```
# Remove outliers
Remove counties that have less than 1000 people, remove counties with a 100% or 0% covid vaccination rate
```{r}
pop_limit = 1000

OneYear <- OneYear%>%
  filter(Covid_Percentage < 100)%>%
  filter(Covid_Percentage > 1)

FirstMonthjoinedPlus_Out <- FirstMonthjoinedPlus %>%
  filter(Census2019_18PlusPop > pop_limit) %>%
  filter(Covid_Percentage < 100) %>%
  filter(Covid_Percentage > 1)

SixMonthsjoinedPlus_Out <- SixMonthsjoinedPlus %>%
  filter(Census2019_18PlusPop > pop_limit) %>%
  filter(Covid_Percentage < 100) %>%
  filter(Covid_Percentage > 1)

oneYearjoinedPlus_Out <- oneYearjoinedPlus %>%
  filter(Census2019_18PlusPop > pop_limit) %>%
  filter(Covid_Percentage < 100) %>%
  filter(Covid_Percentage > 1)

```

# Replot data with outliers removed
Note: Counties with a population below 1000 were removed because of low sample count. A lot of counties have a 0% vaccination rate, particularly in the first 6 months because Covid vaccines were not available. These data points were removed, although they heavily impacted our study as our information may better represent Covid vaccine availability as opposed to Covid Vaccine acceptance.
```{r}
#FirstMonthJoined
#SixMonthsjoined
#OneYearJoined


FirstMonthjoinedPlus_Out %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage)) +
        geom_point()+
        labs(title = "One months after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

 # cor(FirstMonthjoinedPlus$Influenza_Percentage, FirstMonthjoinedPlus$Covid_Percentage)


SixMonthsjoinedPlus_Out %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage))+
        geom_point()+
        labs(title = "Six months after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

##  cor(SixMonthsCors$Influenza_Percentage, SixMonthsCors$Covid_Percentage)

oneYearjoinedPlus_Out %>%
        ggplot(aes(x = Influenza_Percentage, y = Covid_Percentage)) +
        geom_point()+
        labs(title = "One Year after the release of the covid 19 Vaccine") +
        geom_smooth(method = "loess",
              formula = y ~ x)

#OneYearCors<- oneYearjoinedPlus%>%
  #cor(Influenza_Percentage, Covid_Percentage)
```
# Calculating correlation coefficient
The correlation is nearly nonexistent in the first month, however it grows quickly. This trend may continue into later years, however our study only goes one year in. This is likely due to vaccine availability, although it may represent how the willingness to take the covid vaccine was less related to vaccine willingness in general and more of various other factors like public mistrust in the safety of a vaccine produced so quickly. The correlation coefficient however never reaches a value of 0.5 or above, so the correlation isn't strong enough to substantiate any claim regarding covid vaccine willingness.
```{r}

firstMonthCor <- cor(FirstMonthjoinedPlus_Out$Influenza_Percentage, FirstMonthjoinedPlus_Out$Covid_Percentage, use='complete.obs')
paste("First month corrrelation coefficent", firstMonthCor)

SixMonthsCor <- cor(SixMonthsjoinedPlus_Out$Influenza_Percentage, SixMonthsjoinedPlus_Out$Covid_Percentage, use='complete.obs')
paste("Six month corrrelation coefficent", SixMonthsCor)

oneYearCor <- cor(oneYearjoinedPlus_Out$Influenza_Percentage, oneYearjoinedPlus_Out$Covid_Percentage,  use='complete.obs')
paste("One year corrrelation coefficent", oneYearCor)
```
